---
title: "Sce stressed expression" 
author: "h qin"
date: "2016 March 29"
output: html_document
---

Expression data from S. cerevisiae under serial stresses
GSE15936
http://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE15936

See http://www.ncbi.nlm.nih.gov/pubmed/19536156

GPL2529	[Yeast_2] Affymetrix Yeast Genome 2.0 Array

```{r, message=FALSE}
rm(list=ls())
setwd("~/github/GSE15936_yeast_stress")
library(Biobase)
library(GEOquery)
library(limma)
library(foreach)
library(doMC)
```

Load series and platform data from GEO
After retrieve it from NCBI GEO, ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE15nnn/GSE15936/matrix/
Qin saved the data locally for in-class exercises. 

```{r}
 gset <- getGEO(filename="GSE15936_series_matrix.txt.gz")
```

get expression matrix
```{r}
ex <- exprs(gset)
```

Normalize them by total intensity
```{r}
means = apply(ex, 2, mean)
scale = max(means)
for(i in 1:length(ex[1,])){
  ex[,i] = ex[,i] * scale/ means[i]
}
apply( ex, 2, mean) / scale
apply( ex, 2, median) 
boxplot( ex )
```
What is boxplot? See http://tinyurl.com/zw7bd9j

Study the structure of gset
```{r, comment=NA}
str(gset)
gset@phenoData@data
```

Find the experimental design information
```{r, comment=NA}
#gset@phenoData@data
experimental_design = gset@phenoData@data #These should contain experiment meta info
experimental_design[1:3,]
# gset@phenoData@varMetadata
# experimental_design[, "source_name_ch1"][1:10]
experimental_design[5:8,  c("title", "source_name_ch1")]
```

Pick oxidative stress time course
```{r}
experimental_design$title
experimental_design$title[grep("Oxidative stress, t", experimental_design$title)]
selected = c(1, 2, grep("Oxidative stress, t", experimental_design$title))
ex = ex[, selected]
```

Regression for H2O2 time dependent changes 
```{r}
mylevels = c(0,0,15,30,45,90)
names(mylevels) = c("0","0", "15","30","45","90")
my.pvalues = numeric(length=length(ex[,1]))

registerDoMC(cores=4)
#my.pvalues = foreach( i = 1:100, .combine='rbind') %dopar% {
my.pvalues = foreach( i = 1:length(ex[,1]), .combine='rbind') %dopar% {
  m = lm( ex[i,] ~ mylevels )
  sm = summary(m)
  pf(sm$fstatistic[1], sm$fstatistic[2], sm$fstatistic[3], lower.tail = FALSE)
}
row.names(my.pvalues) = row.names(gset@assayData$exprs)
head(my.pvalues)
```

Load NCBI platform annotation
```{r, message=FALSE}
gpl <- annotation(gset)
platf <- getGEO(gpl, AnnotGPL=TRUE)
ncbifd <- data.frame(attr(dataTable(platf), "table"))
ncbifd[1:10,]
```

pick significant genes
```{r}
my.pvalues.BH = p.adjust(my.pvalues, "BH")
names(my.pvalues.BH) = row.names(gset@assayData$exprs)
hist(my.pvalues.BH)
```

```{r, comment=NA}
sig= my.pvalues.BH[my.pvalues.BH < 0.05]
sig= data.frame(sig)
sig$ID = row.names(sig)
#ncbifd[ match(names(sig), as.character( ncbifd$ID) ),  ]
sig2 = merge(sig, ncbifd, by="ID")
sig2
```

